\raggedright
\begin{thebibliography}{bla}

\bibitem{YFullMutationRate} Dmitry Adamov, Vladimir Guryanov,
Sergey Karzhavin, Vladimir Tagankin, Vadim Urasin.
\emph{\href{http://rjgg.molgen.org/index.php/RJGGRE/article/view/151}
{Defining a New Rate Constant for Y-Chromosome SNPs based on Full Sequencing Data}}.
The Russian Journal of Genetic Genealogy
(\foreignlanguage{russian}{Русская версия}),
Vol 6, No 2 (2014)/Vol 7, No 1 (2015).

\bibitem{Newick} James Archie, William H.E. Day, Wayne Maddison,
Christopher Meacham, F. James Rohlf, David Swofford, Joe Felsenstein,
\emph{\href{http://evolution.genetics.washington.edu/phylip/newicktree.html}
{The Newick Tree Format}}.
1986, Date accessed: 2014-08-06.

\bibitem{Trex} Alix Boc, Alpha Boubacar Diallo, Vladimir Makarenkov,
\emph{T-REX: a web server for inferring, validating and
visualizing phylogenetic trees and networks}.
Nucleic Acids Research (2012) 40 (W1): W573-W579.
\href{http://dx.doi.org/10.1093/nar/gks485}{doi: 10.1093/nar/gks485}, 2012. 

\bibitem{Can14} R. A. Canada,
\emph{\href{https://www.familytreedna.com/learn/y-dna-testing/y-str/infinite-allele-palindromic-markers/}
{How does the infinite allele comparison method work for palindromic markers?}}.
Family Tree DNA, 2014, Date accessed: 2014-08-07.

\bibitem{BigYWhitePaper} Family Tree DNA,
\emph{\href{https://www.familytreedna.com/learn/wp-content/uploads/2014/08/BIG_Y_WhitePager.pdf}{Big Y White Paper}}.
Family Tree DNA, August 2014.

\bibitem{Phylip} Joe Felsenstein,
\emph{\href{http://evolution.genetics.washington.edu/phylip.html}
{PHYLIP, a free package of programs for inferring phylogenies.}}.
University of Washington, First version: 1980, Date accessed: 2014-08-06.

\bibitem{Kly09} Anatole A. Klyosov,
\emph{\href{http://www.jogg.info/52/files/Klyosov1.pdf}
{DNA Genealogy, Mutation Rates, and Some Historical
Evidence Written in Y-Chromosome, Part I:  Basic Principles and
the Method}}.
Journal of Genetic Genealogy, 5(2):186-216, 2009.

\bibitem{Kly12} Anatole A. Klyosov,
\emph{Ancient History of the Arbins, Bearers of Haplogroup R1b, from
Central Asia to Europe, 16,000 to 1500 Years before Present}.
Advances in Anthropology, Vol.2, No.2, 87-105,
\href{http://dx.doi.org/10.4236/aa.2012.22010}{doi:10.4236/aa.2012.22010}, 2012.

\bibitem{ISOGGSNPCriteria} ISOGG,
\emph{\href{http://www.isogg.org/tree/ISOGG_SNP_Requirements.html}
{Listing Criteria for SNP Inclusion into the ISOGG Y-DNA Haplogroup Tree - 2015}}.
Date visited: 2015-10-06.

\bibitem{Shl09} Jonathon Shlens,
\emph{\href{http://rieke-server.physiol.washington.edu/People/Fred/Classes/545/shlens-pca.pdf}
{A Tutorial on Principal Component Analysis}}.
Center for Neural Science, New York University,
New York City, 
Systems Neurobiology Laboratory, Salk Insitute for Biological Studies,
La Jolla, 2009.

\bibitem{Wiki-List_of_DYS_markers} Wikipedia,
\emph{\href{http://en.wikipedia.org/wiki/List_of_DYS_markers}
{List of Y-STR markers}}.
Date accessed: 2014-08-15.

\bibitem{Wal02} Bruce Walsh,
\emph{\href{http://nitro.biosci.arizona.edu/ftDNA/models.html}
{Details on the various assumptions used in computing TMRCA}}.
Family Tree DNA, 2002, Date accessed: 2014-08-07.

\bibitem{Wal01} Bruce Walsh,
\emph{\href{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC1461668/}
{Estimating the Time to the Most Recent Common Ancestor for the
Y chromosome or Mitochondrial DNA for a Pair of Individuals}}.
Genetics Society of America, 2001.

\end{thebibliography}
